Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?

### 2016-07-01 text/ms.tex
updating the text

%%\documentclass[12pt, preprint]{aastex}
\documentclass[14pt, preprint]{emulateapj}
\usepackage{microtype}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

@@

\title{Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?}
\author{Mohammadjavad Vakili \altaffilmark{1,$\dagger$}, David W. Hogg, ChangHoon~Hahn, Unspecified contributors}
\altaffiltext{$\dagger$}{\url{mjvakili@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY} 
\begin{abstract}

@@

\section{Introduction}
The assumption that galaxies reside dark matter halos 
is the central tenet of the modern cosmology and 
large-scale structure formation theories. 
One of the most powerful methods for describing 
the galaxy-halo connection is the halo occupation modeling. 
The underlying assumption of the traditional HOD framework is 
that the halo mass alone is sufficient in determining the
population of galaxies inside halos. That is, the statistical properties of 
galaxies is governed by the halo mass. Mathematically, this assumption can be 
written as $P(N_g|M_h,\{x\})=P(N_g|M_h)$ where $\{x\}$ is the set of all 
possible halo properties beyond mass $M_{h}$. HOD provides an analytical prescription for the

Despite this very simplifying assumption, the galaxy-halo connection models such as HOD and CLF have been widely and successfully used in modeling the measurements of a wide range of the statistical properties of galaxies such as . However, the assumption that the statistical 
distribution of galaxies can be determined by halo mass is not a derived characteristic of 
accurate and high resolution hydrodynamic simulations, but rather a theoretical assumption 
that needs to be scrutinized. In recent years, the idea that the

@@

For populating the halos with galaxies, we follow the procedure described in CITE[HEARIN ETAL], CITE[CHANG ETAL]. The central galaxies are assumed to be at the center of the host dark matter halos. We place the satellite galaxies are within the virial radius of the halo following an NFW profile [CITE NFW]. The concentration of the NFW profile is given by the empirical mass-concentration relation provided by [CITE MACCIO]. The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations [CITE MORE]. We refer the readers to [CITE HEARIN 15] for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

\subsubsection{Assembly bias HOD modeling}\label{subsubsec:decorated}
Now let us provide a brief overview of HOD modeling decorated with $\mathtt{Heaviside}$ $\mathtt{Assemblybias}$ introduced in [CITE HEARIN2015]. At a fixed halo mass $M_{\rm h}$, halos are split into two populations: population of halos with the 0.5-percentile of highest concentration, and population of halos with 0.5 percentile of lowest concentration.We call the first population "type 1" halos, and the second population "type 2 halos. Following In the decorated HOD model, the expected number of central and satellite galaxies in the two populations is given by

@@

In this decorated occupation model, the allowable range for the quantities $\delta N_{c,1}$ and $\delta N_{s,1}$ is given by 
\begin{eqnarray}
-\langle N_{c} | M_{h}\rangle \leq \delta N_{c} \leq \langle N_{c} | M_{h}\rangle
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq \delta N_{s} \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, CITE[HEARIN2015] defines the assembly bias parameter $\mathcal{A}$ in the following way:
\begin{eqnarray}
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm max}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} > 0,  \\
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm min}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} < 0,
\end{eqnarray}
where the subscript $\alpha = c , s$ for the centrals and satellites respectively, and $\delta N_{\alpha , 1}^{\rm max}(M_{h})$, $\delta N_{\alpha , 1}^{\rm min}(M_{h})$ are given by Eqs. \ref{eq:cen-bounds} and \ref{eq:sat-bounds}.

@@

\subsection{Model Observables}
\subsubsection{group multiplicity function}
Group multiplicity function $g(N)$ is defined as the abundance of the galaxy groups as a 
function of group richness (the number of galaxies in groups) per richness width, measured in bins of richness. 
Galaxy groups in are found by Friends-of-Friends (hereafter FoF) group finder algorithm [CITE FOF]. 
The FoF group finder algorithm links to galaxy $i$ and galaxy $j$ to each other, if the projected distance between the two galaxies $D_{\perp , i,j}$ is less than a specified projected linking length $b{_\perp}$, and the line-of-sight distance between the two galaxies $D_{\parallel, i,j}$ is less than a specified line-of-sight linking length $b{_\parallel}$. These linking lengths are measured in units of the mean inter-particle distance ($n_{\rm g}^{-1/3}$). In order to be consistent with the [CITE BERLIND] measurements, the linking lengths $b_{\perp}$ and $b_{\parallel}$ are set to to 0.15 and 0.74 respectively.

After populating the dark matter halo catalogs with galaxies, we use the $\mathtt{Halotools}$ implementation of the FoF group-finder algorithm to identify the FoF groups with the mentioned linking lengths. Afterwards, we compute the number density of the galaxy groups in bins of richness to measure the group multiplicity function. Note that $g(N)$ is measured in units of $h^{3}\rm{Mpc}^{-3}$.
\subsubsection{projected two-point correlation function}
The projected two point correlation function $w_{\rm{p}}(r_{\rm{p}})$ can be computed by integrating the 3D redshift space 2PCF $\xi(r_{\rm{p}} , \pi)$ along the line-of-sight
\beq
w_{\rm{p}}(r_{\rm{p}}) = 2 \int_{0}^{\pi_{max}} d\pi \;\xi(r_{\rm{p}} , \pi)
\eeq
For our 2PCF calculations, we use the $w_{\rm p}$ measurement functionality of the fast and publicly available pair-counter code $\mathtt{CorrFunc}$ [CITE CORRFUNC].  To be consistent with the SDSS measurements described in \ref{sec:data}, $w_{\rm{p}}(r_{\rm{p}})$ is obtained by the line-of-sight integration to $\pi_{max}=40 \; h^{-1}\rm{Mpc}$. Note that $w_{\rm{p}}(r_{\rm{p}})$ is measured in units of $h^{-1}\rm{Mpc}$.

@@

For the simulations used in this work, we make use of the Rockstar (cite[Behroozi]) halo catalogs in the $z=0$ Bolshoi-Planck high 
resolution $N$-body simulation CITE[Klypin2014]. This simulation was carried out using the adaptive refinement tree code 
(ART, see cite[Kravstov]) code, following the Planck $\Lambda$CDM cosmological parameters 
$\Omega_{\rm m} = 0.307$, $\Omega_{\rm b} = 0.048$, $\Omega_{\Lambda} = 0.693$, $\sigma_{8} = 0.823$, $n_{\rm s}=0.96$, 
$h=0.678$. The Box size for this $N$-body simulation is 250 $h^{-1} \rm{Mpc}$, the number of simulation particles is 2048$^3$, 
the mass per simulation particle $m_{\rm p}$ is $1.5 \times 10^{8} \; h^{-1} M_{\odot}$, and the gravitational softening length 
$\epsilon$ is 1 $h^{-1} \rm{kpc}$.

@@

The first set of measurements consists of the galaxy number densities $n_{g}$ and the group multiplicity functions $g(N)$ described in [CITE BERLIND 2006]. These measurements are made 
for the three Volume limited samples of the SDSS galaxies with luminosity thresholds of $M_{r}<-18$, $M_{r}<-19$, and $M_{r}<-20$. Galaxy groups in [CITE BERLIND etal] are found by Friends-of-Friends (hereafter FoF) group finder algorithm [CITE FOF]. [CITE BERLIND ETAL] sets $b_{\perp}$ to 0.15 and $b_{\parallel}$ to 0.74 to minimize the discrepancy between the multiplicity function, projected size, and the number of FoF groups and those of the halos in a suite of simulations.

The measurements are accompanied by two sets of uncertainty measurements. The first one is the error on $g(N)$ that stems from the dispersion between $g(N)$ measurements of different mock catalogs. The second source of error is the Poisson error that results from the number of groups in each richness bin. Finally, these two uncertainties are added in quadrature 
to estimate the total error on [CITE BERLIND ETAL] $g(N)$ measurements.

@@

The second set consists of the number density $n_{g}$ and the projected 2PCF $w_{\rm{p}}(r_{\rm{p}})$. The project 2PCFs are measured in 12 logarithmic $r_{\rm p}$ bins (in units of $h^{-1}\rm{Mpc}$) of width $\Delta \log(r_{\rm p}) = 0.2$, starting from $r_{\rm{p}} = 0.1 \; h^{-1}\rm{Mpc}$. These measurements are accompanied by their corresponding covariance matrices constructed using 400 jackknife sub-samples of the data. The advantage of using these measurements is that the effects of fiber collision systematics on the two-point statistics are corrected for, and therefore, they provide very accurate small scale clustering measurements [see CITE GUO 2015]. As [CITE HEARIN2015] demonstrates the satellite assembly bias $\mathcal{A}_{\rm sat}$ can have 10 $\%$ level impact on the small scale galaxy clustering.

@@

Now we need to specify the likelihood function $p(d|\theta)$. In the case where the observables are the number density and the GMF are used the negative log-likelihood is given by
\begin{eqnarray}
-2\ln p(d|\theta) &=& \frac{[n^{\rm data}_{g}-n^{\rm model}_{g}]^{2}}{\sigma_{n}^{2}} \nonumber \\ &+& \sum_{N} \frac{[ g^{\rm data}(N)-g^{\rm model}(N)]^{2}}{\sigma^{2}_{N}} \; + \; \rm{const.},
\label{eq:lnlike_group}
\end{eqnarray}
where the summation is over the group richness bins, and $\sigma_{N}$ is the uncertainty associated with the richness bin $N$. In the case, where the number density and 2PCF are used, the negative log-likelihood is given by
\begin{eqnarray}
-2\ln p(d|\theta) &=& \frac{[n^{\rm data}_{g}-n^{\rm model}_{g}]^{2}}{\sigma_{n}^{2}} \nonumber \\  
&+&  \sum_{\rm p} \Delta \wpp^{\rm T}C^{-1}\Delta \wpp \; + \; \rm{const.},
\label{eq:lnlike_wp}
\end{eqnarray}
where $\Delta \wpp(\rpp) = \wpp^{\rm data}(\rpp)-\wpp^{\rm model}(\rpp)$, and  $C^{-1}$ is inverse of the estimate of the jackknife covariance matrix provided by [CITE GUO2015]. 
Another important ingredient of our analysis is specification of the prior probabilities over the parameters of the two halo occupation models. For both model, we use uniform flat priors for all the parameters. The prior ranges are specified in the Table \ref{tab:prior}.

@@

For sampling from the posterior probability, given the likelihood function (see Eqs. \ref{eq:lnlike_group}, \ref{eq:lnlike_wp}) and the prior probability distributions (see Table \ref{tab:prior}), we use the affine-invariant ensemble MCMC sampler $\mathtt{emcee}$ (CITE[FORMAN-MACKEY]). In particular, we run the $\mathtt{emcee}$ code with 140 walkers and we run the chains for at least 15000 iterations until the convergence of the MCMC chains is ensured. For testing the convergence of the chains, we perform Gelman-Rubin convergence test [CITE GELMANANDRUBIN and the other paper].

@@

There are various methods for model comparison. In this work we make use of two model comparison criteria: \emph{Akaike Information Criterion} (AIC, see CITE ... ), and \emph{Bayesian Information Criteria} (BIC, see CITE ....). AIC and BIC are two widely used methods for assessing the relative quality of models given a set of observations.
Suppose that $\mathcal{L}_{\rm max}$ is the maximum likelihood of the data given a model, $N_{\rm par}$ is the number of parameters of a given model, and $N_{\rm data}$ is the number of data points. Then we have

\begin{eqnarray}
AIC&=& -2\;\ln \mathcal{L}_{\rm max} + 2N_{\rm par} + 2\;\frac{N_{\rm par}(N_{\rm par}+1)}{N_{\rm data}-N_{\rm par}-1} \; , \label{eq:aic}\\
BIC&=& -2\;\ln \mathcal{L}_{\rm max} + N_{\rm par}\;\ln N_{\rm data} \; . \label{eq:bic}
\end{eqnarray}
Models that deliver lower values of information criteria are more favorable. In Eqs. (\ref{eq:aic}) and (\ref{eq:bic}) $N_{\rm par}$ is 5 when the standard HOD model is used, and 7 when the Heaviside assembly bias model is used. When the observables are $\ngal$ and $\wpp$, $N_{\rm data}$ is 13 for all luminosity thresholds, and when the observables are $\ngal$ and $\gmf$, $N_{\rm data}$ is 16 for the $\mr < -18$ luminosity threshold, 23 for the $\mr < -19$ luminosity threshold, and 25 for the $\mr < -20$ luminosity threshold.

@@

\begin{figure*}
\includegraphics[width=\textwidth]{paper_abhod_ncengmf.pdf}
\includegraphics[width=\textwidth]{paper_abhod_nsatgmf.pdf}
\includegraphics[width=\textwidth]{paper_abvshod_gmf.pdf}

\caption{We demonstrate the posterior predictions for the expected occupation statistics of galaxies. The shaded regions mark the 68 $\%$ confidence region of the posterior predictions. The observables used in fitting are $\bar{n}$ and $\gmf$. $\bf{Top \; panel}$: posterior predictions for the expected number of centrals as a function of halos for the two population of halos: $type$-1 (high concentration, sown with blue) and $type$-2 (low concentration, shown with yellow) halos. }
\label{fig:gmf-hod-predictions}
\end{figure*}

>>>>

%%\documentclass[12pt, preprint]{aastex}
\documentclass[12pt, preprint]{emulateapj}
\usepackage{microtype}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

@@

\title{Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?}
\author{Mohammadjavad Vakili  \altaffilmark{1,$\dagger$}, ChangHoon~Hahn \altaffilmark{1}, Unspecified contributors \altaffilmark{1}}
\altaffiltext{$\dagger$}{\url{mjvakili@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY} 
\begin{abstract}

@@

\section{Introduction}
The assumption that galaxies reside dark matter halos 
is the central tenet of the modern cosmology and 
large-scale structure formation theories. The study of galaxy-halo connection is an essential tool for constraining the cosmology via small scale clustering and galaxy formation physics. One of the most powerful methods for describing the galaxy-halo connection is the halo occupation distribution (HOD, see \citealt{seljak2000, berlind_weinberg2002, scoccimarro2001, zheng2005, zheng2007}) and conditional luminosity function (CLF, see \citealt{clf_vdBosch}).

HOD provides an analytical prescription for the statistical properties of galaxies within dark matter halos by specifying a probability distribution function $P(N|x)$ where $N$ is the number of galaxies and $x$ is a property of the dark matter halo. The underlying assumption of the standard HOD framework is 
that the halo mass alone is sufficient in determining the
population of galaxies inside halo. That is, the statistical properties of 
galaxies is governed by the halo mass. Mathematically, this assumption can be 
written as $P(N_g|M_h,\{x\})=P(N_g|M_h)$ where $\{x\}$ is the set of all 
possible halo properties beyond mass $M_{h}$. In a similar fashion, CLF ties the luminosity of galaxies to the host halo mass.

Despite this very simplifying assumption, the galaxy-halo connection models based on HOD and CLF have been used in fitting the measurements of a wide range of statistics such as correlation function of galaxies and galaxy-galaxy lensing. HOD has been used in constraining the cosmological parameters through modeling the galaxy two-point correlation function (hereafter 2PCF) (\citealt{abazajian2005}), combination of 2PCF with mass-to-light ratio of galaxies (\citealt{tinker05}), redshift space distortions (\citealt{tinker_rsd2007}), mass-to-number ratio of galaxy clusters (\citealt{tinker2012}) galaxy-galaxy lensing in SDSS main sample of galaxies (\citealt{vdb03,cacciato13,more13,vdb13}), and galaxy-galaxy lensing in the BOSS galaxies (\citealt{miyatake15,more15}). It has also been used in galaxy evolution studies (\citealt{conroy09,leauthaud12,behroozi13}).

It has been shown by the $N$-body simulations that the halo clustering is correlated with the formation history of halos. That is, at a fixed halo mass, the halo bias is correlated with properties of halos beyond mass such as the concentration, formation time, spin, and etc. This phenomenon is known as the halo assembly bias (see \citealt{sheth2004,gao2005, harker2006, weschler2006, gao2007,croton2007,wang2007,angulo2008,dalal2008,li2008,sunayama2016}). From the observational standpoint, this phenomenon has been confirmed
(\citealt{miyatake2016,more2016}).

If the galaxy properties are tied to the formation history of their host halos, we expect the occupation statistics of galaxies to be coupled to the properties as the concentration of halos, environment, and etc. This is called galaxy assembly bias. In the context of the semi-analytic models of galaxy formation, \citealt{croton2007} shows that

However, the assumption that the statistical 
distribution of galaxies can be determined by halo mass is not a derived characteristic of accurate and high resolution hydrodynamic simulations, but rather a theoretical assumption that needs to be scrutinized. In recent years, the idea that the

@@

For populating the halos with galaxies, we follow the procedure described in \citealt{decorated}, \todo{HAHN+VAKILI 16 preparation}. The central galaxies are assumed to be at the center of the host dark matter halos. We place the satellite galaxies are within the virial radius of the halo following an NFW profile (\citealt{nfw}). The concentration of the NFW profile is given by the empirical mass-concentration relation provided by (\citealt{nfw_c(M)}). The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations (\citealt{more2010}). We refer the readers to \citealt{decorated} for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

As described in \citealt{decorated} and \todo{HAHN+VAKILI in prep}, this approach makes no appeal to the fitting functions used in the analytical calculation of the 2PCF. The accuracy of these fitting functions is limited (\citealt{tinker08,tinker10,watson13}). This approach also does not face the known issues of properly treatment of halo exclusion and scale-dependent bias that can lead to significant inaccuracies (\citealt{vdb13}).

Another added advantage of this forward modeling approach is that it permits us to take advantage of observables beyond the traditionally used 2PCF. One of these observables is the group multiplicity function which characterizes the abundance of galaxy groups. There is no fitting function for calculation of this observable, and therefore, our approach of direct population of dark matter halos allows us to use these unconventional measurements for constraining the parameters governing the galaxy-halo collection.

\subsubsection{Assembly bias HOD modeling}\label{subsubsec:decorated}
Now let us provide a brief overview of HOD modeling decorated with $\mathtt{Heaviside}$ $\mathtt{Assemblybias}$ introduced in \citealt{decorated}. At a fixed halo mass $M_{\rm h}$, halos are split into two populations: population of halos with the 0.5-percentile of highest concentration, and population of halos with 0.5 percentile of lowest concentration.We call the first population "type-1" halos, and the second population "type-2 halos. In the decorated HOD model, the expected number of central and satellite galaxies in the two populations is given by

@@

In this decorated occupation model, the allowable range for the quantities $\delta N_{c,1}$ and $\delta N_{s,1}$ is given by 
\begin{eqnarray}
-\langle N_{c} | M_{h}\rangle \leq \delta N_{c} \leq \langle N_{c} | M_{h}\rangle
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq \delta N_{s} \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, the assembly bias parameter $\mathcal{A}$ is defined in the following way:
\begin{eqnarray}
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm max}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} > 0,  \\
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm min}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} < 0,
\end{eqnarray}
where the subscript $\alpha = c , s$ for the centrals and satellites respectively, and $\delta N_{\alpha , 1}^{\rm max}(M_{h})$, $\delta N_{\alpha , 1}^{\rm min}(M_{h})$ are given by Eqs. \ref{eq:cen-bounds} and \ref{eq:sat-bounds}.

@@

\subsection{Model Observables}
\subsubsection{group multiplicity function}
Group multiplicity function $g(N)$ is defined as the abundance of the galaxy groups as a 
function of group richness (the number of galaxies in groups) per richness width, measured in bins of richness. 
Galaxy groups in are found by Friends-of-Friends (hereafter FoF) group finder algorithm (\citealt{fof}). 
The FoF group finder algorithm links to galaxy $i$ and galaxy $j$ to each other, if the projected distance between the two galaxies $D_{\perp , i,j}$ is less than a specified projected linking length $b{_\perp}$, and the line-of-sight distance between the two galaxies $D_{\parallel, i,j}$ is less than a specified line-of-sight linking length $b{_\parallel}$. These linking lengths are measured in units of the mean inter-particle distance ($n_{\rm g}^{-1/3}$). In order to be consistent with the \citealt{berlind2006} measurements described in section \ref{sec:data}, the linking lengths $b_{\perp}$ and $b_{\parallel}$ are set to to 0.15 and 0.74 respectively.

After populating the dark matter halo catalogs with galaxies, we use the $\mathtt{Halotools}$ implementation of the FoF group-finder algorithm to identify the FoF groups with the mentioned linking lengths. Afterwards, we compute the number density of the galaxy groups in bins of richness to measure the group multiplicity function. Note that $g(N)$ is measured in units of $h^{3}\rm{Mpc}^{-3}$.
\subsubsection{projected two-point correlation function}
The projected two point correlation function $w_{\rm{p}}(r_{\rm{p}})$ can be computed by integrating the 3D redshift space 2PCF $\xi(r_{\rm{p}} , \pi)$ along the line-of-sight
\beq
w_{\rm{p}}(r_{\rm{p}}) = 2 \int_{0}^{\pi_{max}} d\pi \;\xi(r_{\rm{p}} , \pi)
\eeq
For our 2PCF calculations, we use the $w_{\rm p}$ measurement functionality of the fast and publicly available pair-counter code $\mathtt{CorrFunc}$ (\citealt{corrfunc}).  To be consistent with the SDSS measurements described in \ref{sec:data}, $w_{\rm{p}}(r_{\rm{p}})$ is obtained by the line-of-sight integration to $\pi_{max}=40 \; h^{-1}\rm{Mpc}$. Note that $w_{\rm{p}}(r_{\rm{p}})$ is measured in units of $h^{-1}\rm{Mpc}$.

@@

For the simulations used in this work, we make use of the Rockstar (\citealt{rockstar}) halo catalogs in the $z=0$ Bolshoi-Planck high 
resolution $N$-body simulation (\citealt{bolshoi}). This simulation was carried out using the adaptive refinement tree code 
(ART, see \citealt{art}) code, following the Planck $\Lambda$CDM cosmological parameters 
$\Omega_{\rm m} = 0.307$, $\Omega_{\rm b} = 0.048$, $\Omega_{\Lambda} = 0.693$, $\sigma_{8} = 0.823$, $n_{\rm s}=0.96$, 
$h=0.678$. The Box size for this $N$-body simulation is 250 $h^{-1} \rm{Mpc}$, the number of simulation particles is 2048$^3$, 
the mass per simulation particle $m_{\rm p}$ is $1.5 \times 10^{8} \; h^{-1} M_{\odot}$, and the gravitational softening length 
$\epsilon$ is 1 $h^{-1} \rm{kpc}$.

@@

The first set of measurements consists of the galaxy number densities $n_{g}$ and the group multiplicity functions $g(N)$ described in \citealt{berlind2006}. These measurements are made in three Volume-limited samples of the SDSS galaxies with luminosity thresholds of $M_{r}<-18$, $M_{r}<-19$, and $M_{r}<-20$. These galaxy groups are found by Friends-of-Friends (hereafter FoF) group finder algorithm (\citealt{fof}). \citealt{berlind2006} sets $b_{\perp}$ to 0.15 and $b_{\parallel}$ to 0.74 to minimize the discrepancy between the multiplicity function, projected size, and the number of FoF groups and those of the halos in a suite of $N$-body simulations.

The measurements are accompanied by two sets of uncertainty measurements. The first one is the error on $g(N)$ that stems from the dispersion between $g(N)$ measurements of different mock catalogs. The second source of error is the Poisson error that results from the number of groups in each richness bin. Finally, these two uncertainties are added in quadrature 
to estimate the total error on these $g(N)$ measurements.

@@

The second set consists of the number density $n_{g}$ and the projected 2PCF $w_{\rm{p}}(r_{\rm{p}})$ made by \citealt{guo2015} for Volume-limited sample of galaxies in SDSS DR7 galaxy catalog. The project 2PCFs are measured in 12 logarithmic $r_{\rm p}$ bins (in units of $h^{-1}\rm{Mpc}$) of width $\Delta \log(r_{\rm p}) = 0.2$, starting from $r_{\rm{p}} = 0.1 \; h^{-1}\rm{Mpc}$. These measurements are accompanied by their corresponding covariance matrices constructed using 400 jackknife sub-samples of the data. The advantage of using these measurements is that the effects of fiber collision systematics on the two-point statistics are corrected for, and therefore, these measurements provide very accurate small scale clustering measurements (\citealt{guo2012,guo2015}). As \citealt{hearin15} demonstrates, the satellite assembly bias $\mathcal{A}_{\rm sat}$ can lead to 10 $\%$ level impact on the small scale galaxy clustering.

@@

Now we need to specify the likelihood function $p(d|\theta)$. In the case where the observables are the number density and the GMF are used the negative log-likelihood is given by
\begin{eqnarray}
-2\ln p(d|\theta) &=& \frac{[n^{\rm data}_{g}-n^{\rm model}_{g}]^{2}}{\sigma_{n}^{2}} \nonumber \\ &+& \sum_{N} \frac{[ g^{\rm data}(N)-g^{\rm model}(N)]^{2}}{\sigma^{2}_{N}} \; + \; \rm{const.},
\label{eq:lnlike_group}
\end{eqnarray}
where the summation is over the group richness bins, and $\sigma_{N}$ is the uncertainty associated with the richness bin $N$. In the case, where the number density and 2PCF are used, the negative log-likelihood is given by
\begin{eqnarray}
-2\ln p(d|\theta) &=& \frac{[n^{\rm data}_{g}-n^{\rm model}_{g}]^{2}}{\sigma_{n}^{2}} \nonumber \\  
&+&  \sum_{\rm p} \Delta \wpp^{\rm T}C^{-1}\Delta \wpp \; + \; \rm{const.},
\label{eq:lnlike_wp}
\end{eqnarray}
where $\Delta \wpp(\rpp) = \wpp^{\rm data}(\rpp)-\wpp^{\rm model}(\rpp)$, and  $C^{-1}$ is inverse of the estimate of the jackknife covariance matrix provided by \citealt{guo15}. 
Another important ingredient of our analysis is specification of the prior probabilities over the parameters of the two halo occupation models. For both model, we use uniform flat priors for all the parameters. The prior ranges are specified in the Table \ref{tab:prior}.

@@

For sampling from the posterior probability, given the likelihood function (see Eqs. \ref{eq:lnlike_group}, \ref{eq:lnlike_wp}) and the prior probability distributions (see Table \ref{tab:prior}), we use the affine-invariant ensemble MCMC sampler (\citealt{goodmanweare}) and its implementation $\mathtt{emcee}$ (\citealt{emcee}). In particular, we run the $\mathtt{emcee}$ code with 140 walkers and we run the chains for at least 15000 iterations until the convergence of the MCMC chains is ensured. For testing the convergence of the chains, we perform Gelman-Rubin convergence test (\citealt{grtest}).

@@

There are various methods for model comparison. In this work we make use of two model comparison criteria: \emph{Akaike Information Criterion} (AIC, \citealt{aic} , see \citealt{gelmanic} for detailed discussion in AIC), and \emph{Bayesian Information Criteria} (BIC, \citealt{bic}). AIC and BIC are two widely used methods for assessing the relative quality of models given a set of observations.
Suppose that $\mathcal{L}_{\rm max}$ is the maximum likelihood of the data given a model, $N_{\rm par}$ is the number of parameters of a given model, and $N_{\rm data}$ is the number of data points. Then we have

\begin{eqnarray}
\rm{AIC}&=& -2\;\ln \mathcal{L}_{\rm max} + 2N_{\rm par} \nonumber \\
         &+& 2\;\frac{N_{\rm par}(N_{\rm par}+1)}{N_{\rm data}-N_{\rm par}-1} \; , \label{eq:aic}\\
\rm{BIC}&=& -2\;\ln \mathcal{L}_{\rm max} + N_{\rm par}\;\ln N_{\rm data} \; . \label{eq:bic}
\end{eqnarray}
Models that deliver lower values of information criteria are more favorable. In Eqs. (\ref{eq:aic}) and (\ref{eq:bic}) $N_{\rm par}$ is 5 when the standard HOD model is used, and 7 when the Heaviside assembly bias model is used. When the observables are $\ngal$ and $\wpp$, $N_{\rm data}$ is 13 for all luminosity thresholds, and when the observables are $\ngal$ and $\gmf$, $N_{\rm data}$ is 16 for the $\mr < -18$ luminosity threshold, 23 for the $\mr < -19$ luminosity threshold, and 25 for the $\mr < -20$ luminosity threshold.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.5]{paper_abhod_ncengmf.pdf}\\
\includegraphics[scale=0.5]{paper_abhod_nsatgmf.pdf}\\
\includegraphics[scale=0.5]{paper_abvshod_gmf.pdf}

\caption{We present the posterior predictions for the expected number of galaxies as a function of host halo mass in three luminosity threshold samples: $\mr<-18$, $\mr<-19$, and $\mr<-20$. The shaded regions mark the 68 $\%$ confidence region of the posterior predictions. The observables used in fitting are $\bar{n}$ and $\gmf$. $\bf{Top \; panel}$: posterior predictions for the expected number of centrals as a function of host halo mass for the two population of halos: $type$-1 (high-$c$, shown with blue) and $type$-2 (low-$c$, shown with yellow) halos. The expected number of centrals in $type$-1 and $type$-2 halos at a fixed halo mass are in agreement. $\bf{Middle \; panel}$: posterior predictions for the expected number of satellites as a function of host halo mass for the two population of halos. The expected number of satellites in two population of halos at a fixed halo mass are in agreement.$\bf{Bottom \; panel}$: posterior predictions for the expected number of galaxies as a function of host halo mass for two models: the standard HOD (shown in red), and decorated HOD model (shown in blue). The prediction of the two models for the expected number of galaxies as a function of host halo mass are in agreement.}
\label{fig:gmf-hod-predictions}
\end{center}
\end{figure*}
